systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2005-11-02 paper/trunk/systemfit.tex
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%% an abstract and keywords
\Abstract{ Many statistical analysis are based on models that consist
  of  more than one equation.  These equations should be fitted
  simultaneously rather than equation  by equation.  The package
  \pkg{systemfit} provides the capability to estimate systems  of
  linear and non-linear equations in \proglang{R}.
}

\Keywords{R, simultaneous equation system, seemingly unrelated regression}

@@

\input{systemfit_intro.tex}
\input{systemfit_estimation.tex}
\input{systemfit_restrictions.tex}
\input{systemfit_other.tex}
\input{systemfit_usage.tex}
\input{systemfit_summary.tex}

\section{CVS Revision Log}
\begin{verbatim}
$Log$
Revision 1.12  2005/09/26 11:24:05  henningsena
some minor changes

>>>>

%% an abstract and keywords
\Abstract{Many statistical analysis are based on sets of models that
  consist of  more than one equation which contain relationships among
  the various equations.  Ignoring these relationships can result in
  biased and possibly inconsistent parameter estimates. These sets of
  equations should be fitted simultaneously rather than equation  by
  equation.  The package \pkg{systemfit} provides the capability to
  estimate systems of linear and non-linear equations, using limited
  and full-information methods within the \proglang{R} programming
  environment.  }

\Keywords{R, simultaneous equation system, seemingly unrelated
  regression, two-stage least squares, three-stage least squares}

@@

\input{systemfit_intro.tex}
\input{systemfit_estimation.tex}
\input{systemfit_restrictions.tex}
\input{systemfit_usage.tex}
\input{systemfit_other.tex}
\input{systemfit_summary.tex}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Estimating nonlinear equation systems}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The \pkg{systemfit} package also contains a method for fitting
% nonlinear systems of equations.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Nonlinear parameter estimation}
% \label{sec:nonlinear_estimation}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% A system of nonlinear equations can be written as:

% \begin{equation}
%   \label{eq:non_linear_eq_1}
%   \epsilon_{t} = q( y_t, x_t, \beta )
% \end{equation}
% %%% It is not clear what these variable are. 
% %%% Adding an equation representing only a single equation
% %%% would make things clearer.

% \noindent and

% \begin{equation}
%   \label{eq:non_linear_eq_2}
%   z_{t} = Z( x_t )
% \end{equation}

% where $\epsilon_{t}$ are the residuals from the y observations and
% $z_{t}$ are the function evaluated at the parameter estimates.

% \textbf{you really need to check this!!!}

% The objective functions for the methods are:

% \begin{center}
% \begin{tabular}{|l|c|c|c|} \hline
%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
% \end{tabular}
% \end{center}

% where, $r$ is a column vector for the residuals for each equation,
% %% what is the difference to \epsilon?
% $X$ is matrix of the partial derivatives of the dependent variable 
% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
% %% is this correct?
% %% in the linear section X is used for the regressors, 
% %% thus we should use a different variable name here.
% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
% matrix of the instrument variables, and $I$ is an $n \times n $
% identity matrix and $S$ is the estimated variance-covariance matrix between the
% equations

% \begin{equation}
%   \label{eq:non-linear_varcov}
%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
% \end{equation}

% \textbf{The residual variance-covariance matrix can be calculated in 
% different ways (section~\ref{sec:residcov}). 
% It should be relatively easy to implement this also in nlsystemfit(), 
% e.g. the function 'calcRCov' in systemfit() could be moved outside
% systemfit() that it can be used also by nlsystemfit().}

% \textbf{You need to clear this up.}

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % \subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % \code {nlsystemfit} fits a set of structural nonlinear equations using
% % Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
% % Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
% % the objective functions described in section
% % \ref{sec:nonlinear_estimation}.

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % \subsubsection{Standard usage}

% % Similar to calling the \code{systemfit} function, \code{nlsystemfit}
% % is called with a minimum of three arguments,

% % \code{
% % R> nlsystemfit( method, eqns, start )
% % }

% % where \code{method} is one of the following estimation methods: "OLS",
% % "SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
% % to those described for \code{systemfit}, and \code{start} is a list of
% % starting values of the parameter estimates.

% % % The mandatory argument \code{eqns} is a list of the equations  to
% % % estimate.  Each equation is a standard formula in \proglang{R}.  It
% % % starts with a dependent variable on the left hand side.  After a tilde
% % % ($~$) the regressors are listed  \footnote{For Details see the
% % %   \proglang{R} help files to \code{formula}}.

% % This is now demonstrated using an example: \\

% % % \code{
% % % R> library( systemfit ) \\
% % % R> data( ppine ) \\
% % % R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
% % % R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
% % % R> labels <- list( "height.growth", "diameter.growth" )\\
% % % R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
% % % R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
% % % R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
% % % R> model <- list( hg.formula, dg.formula )\\
% % % R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
% % % R+   eqnlabels=labels, inst=inst ) \\
% % % R> summary( model.3sls ) \\
% % % }

% % The nlsystemfit function relies on \code{nlm} to perform the
% % minimization of the objective functions and the \code{qr} set of
% % functions. If the user does not have a set of estimates for the
% % initial parameters, it is suggested using linearized forms in one of
% % the linear methods, or simply using \code{nls} to obtain estimates for
% % the equations.

% % The outputs are similar to those for a systemfit object.

\section{CVS Revision Log}
\begin{verbatim}
$Log$
Revision 1.15  2005/11/02 18:15:40  hamannj
small changes

Revision 1.14  2005/10/26 21:16:53  hamannj
checked out and updated local copy before edits

Revision 1.13  2005/09/26 20:34:12  henningsena
using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)

Revision 1.12  2005/09/26 11:24:05  henningsena
some minor changes

### 2005-11-02 paper/trunk/systemfit_estimation.tex
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary least squares (OLS)}

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X'X \right)^{-1} X'y
\end{equation}

These estimates are efficient only if the disturbance terms are not 
contemporaneously correlated, which means 
\begin{equation}
   \sigma_{ij} = 0 \; \forall \; i \neq j
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
   \label{eq:OLSvcm2}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted least squares (WLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seemingly unrelated regression (SUR)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two-stage least squares (2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted two-stage least squares (W2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three-stage least squares (3SLS)}

@@

>>>>

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = \mathbf{X}_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
\end{equation}
where $y_i$ is a vector of the dependent variable, $\mathbf{X}_i$ is a
matrix of the exogenous variables, $\beta_i$ is the coefficient vector
and $u_i$ is a vector of the disturbance terms of the $i$th equation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary Least Squares (OLS)}

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by

\begin{equation}
   \widehat{\beta} = \left( X'X \right)^{-1} X'y
\end{equation}

These estimates are efficient only if the disturbance terms are not 
contemporaneously correlated, which means

\begin{equation}
   \sigma_{ij} = 0 \; \forall \; i \neq j
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is

\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
\end{equation}

with $\sigma^2 = E \left( u' u \right)$.  This assumes that the
disturbances of all equations have the same variance.

If the disturbance terms of the individual equations  are allowed to
have different variances,  the variance-covariance matrix of the
estimated parameters is

\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
   \label{eq:OLSvcm2}
\end{equation}

\noindent with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and $\sigma_{ii} = E \left(
  u_i' u_i \right)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted Least Squares (WLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seemingly Unrelated Regression (SUR)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two-Stage Least Squares (2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted Two-Stage Least Squares (W2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three-Stage Least Squares (3SLS)}

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End:

### 2005-11-02 paper/trunk/systemfit_intro.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

Many theoretical models consist of more than one equation. If these
models are econometrically estimated, the equations generally should
not be estimated individually, but all together, because the
disturbance terms of these equations are likely contemporaneously
correlated. Estimating all equations simultaneously and taking the
covariance structure of the residuals into account leads to efficient
estimates, while ignoring this correlation leads to an inefficient
parameter estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation
approach. Especially the economic theory suggests many cross-equation
restrictions.

The \pkg{systemfit} package provides the capability to estimate
linear and non-linear equation systems in \proglang{R}.
Although linear and non-linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g. \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
using \pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control many important details of the
estimation procedure that are generally not available in other software
packages.
On the other hand these user options have reasonable defaults that also
beginners can easily use the provided functions.

\textbf{I'll have to update this section to match the recent changes. JDH.}

This paper is organized as follows: In section~\ref{sec:Estimation} we
introduce the mathematics of estimating equation systems.
Section~\ref{sec:Restrictions} shows how linear restrictions can be
imposed.
All other relevant issues are discussed in section~\ref{sec:Other}.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit}, especially how the features presented in the previous
sections can be used.
Finally, a summary and an outlook are given in
section~\ref{sec:Summmary}.

>>>>

Many theoretical models consist of more than one equation. If these
models are econometrically estimated, it may not be appropriate to
estimate the equations individually, but all together, because the
disturbance terms of these equations are likely to be
contemporaneously correlated \citep{theil71} (make sure this is the
correct paper). Estimating all equations simultaneously, taking the
covariance structure of the residuals into account, leads to efficient
estimates. Ignoring this correlation leads to an inefficient parameter
estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation approach
especially the economic theory suggests many cross-equation
restrictions \textbf{(is this statement really true?)}.

The \pkg{systemfit} package provides the capability to estimate linear
and non-linear equation systems in \proglang{R}.  Although linear and
non-linear equation systems can be estimated with several other
statistical and econometric software packages (e.g. \proglang{SAS},
\proglang{EViews}, \proglang{TSP}), \pkg{systemfit} has several
advantages.  First, all estimation procedures are publicly available
in the source code.  Second, the estimation algorithms can be easily
modified to meet specific requirements.  Third, the (advanced) user
can control many important details of the estimation procedure that
are generally not available in other software packages.  On the other
hand, these user options have reasonable defaults that also beginners
can easily use the provided functions.

This paper is organized as follows: In section~\ref{sec:Estimation} we
introduce the mathematics of estimating equation systems.
Section~\ref{sec:Restrictions} shows how linear restrictions can be
imposed.  Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit}, especially how the features presented in the previous
sections can be used. All other relevant issues are discussed in
section~\ref{sec:Other}. Finally, a summary and outlook are presented
in section~\ref{sec:Summmary}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End:

### 2005-11-02 paper/trunk/systemfit_other.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31
using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@576 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

A further approach to obtain the estimated residual variance-covariance
matrix is \citep[p. 309]{zellner62c}
\begin{eqnarray}
   \widehat{\sigma}_{ij} & = & 
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
   & = &
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
   \label{eq:rcov2}
\end{eqnarray} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, "it is doubtful whether [this formula] is really superior to 
[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.

@@

There exist different approaches to determine the degrees of
freedom for testing the estimated parameters with a t-test.

@@

@@

\textbf{Man, I really have to clean this up.}

Hausman \citep{hausman1978} developed a test have power against
alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
under misspecification.

The Hausman result if the difference between the covariance of an
efficient and the covariance of the inefficient estimator is zero, the
model is correctly specified.

@@

Since the models that are fit using the \pkg{systemfit} package are
typically random effects models, the effects may be inconsistent
between the introduced variables and the regressors.

>>>>

A further approach to obtain the estimated residual variance-covariance
matrix is \citep[p. 309]{zellner62c}
\begin{eqnarray}
   \widehat{\sigma}_{ij} & = & 
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
   & = &
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
   \label{eq:rcov2}
\end{eqnarray} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, °it is doubtful whether [this formula] is really superior to 
[(\ref{eq:rcov0})]° \citep[p.322]{theil71}.

@@

Different approaches to determine the degrees of freedom for testing
the estimated parameters with a t-test \textbf{(cite this one)}.

\textbf{Arne, do you have a citation for this?}

The \code{probdfsys} can be used to set the degrees of freedom when
calculating the probability values for the t-test for individual
parameters. This argument can set the degrees of freedom for the
entire system instead of using the degrees of freedom for each of the
individual equations.

%    \item{probdfsys}{use the degrees of freedom of the whole system
%       (in place of the degrees of freedom of the single equation)
%       to calculate prob values for the t-test of individual parameters.}

@@

Currently, the \code{systemfit} package can perform a Likelihood Ratio
test to test linear restrictions when estimating SUR equation systems.

This test calculates the likelihood ratio value by calculating the
estimated residual covariance matrix $hat{S}$ by the maximum likelihood
formula

\begin{equation}
hat{s}_{ij} = \frac{\hat{e}_i^{'} \hat{e}_j}{T} 
\end{equation}
where $T$ is vector of ones.

% (no matter which formula is used in the estimation).

You should really get some of the text from your stats book.

@@

Hausman \citep{hausman1978} developed a test have power against
alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
under misspecification. The resulting test statistic, distributed as a
$\chi^2$, is the difference between the covariance of an efficient and
the covariance of the inefficient estimator is zero, if the model is
correctly specified.

@@

% Since the models that are fit using the \pkg{systemfit} package are
% typically random effects models, the effects may be inconsistent
% between the introduced variables and the regressors.

% results2sls  result of a 2SLS (limited information) estimation returned by systemfit. 
% results3sls  result of a 3SLS (full information) estimation returned by systemfit.

The null hypotheses of the test is that all exogenous variables are
uncorrelated with all disturbance terms. Under this hypothesis both
the 2SLS and the 3SLS estimator are consistent but only the 3SLS
estimator is (asymptotically) efficient. Under the alternative
hypothesis the 2SLS estimator is consistent but the 3SLS estimator is
inconsistent.

The Hausman test statistic is,

\begin{equation}
m = ( b_2 - b_3 )^{'} ( V_2 - V_3 ) ( b_2 - b_3 ) 
\label{eq:hausman}
\end{equation}
where $b_2$ and $V_2$ are the estimated coefficients and the
variance-covariance matrix from a 2SLS estimation and $b_3$ and $V_3$
are the estimated coefficients and their variance covariance-matrix
from a 3SLS estimation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonlinear estimation}

As linear systems of equations can be estimated using
\code{systemfit}, nonlinear systems of equations can be estimated
using a similar function \code{nlsystemfit}. The usage for the
nonlinear estimation function is similar to \code{systemfit} and can
estimate system of equations using OLS, SUR, 2SLS and 3SLS methods.

\begin{equation}
  \label{eq:non_linear_eq_1}
  \epsilon_{t} = q( y_t, x_t, \beta )
\end{equation}

\noindent and

\begin{equation}
  \label{eq:non_linear_eq_2}
  z_{t} = Z( x_t )
\end{equation}
where $\epsilon_{t}$ are the residuals from the $y_t$ observations,
$x_t$ is a matrix of independent variables, $\beta$ is a vector is
parameter estimates and $z_{t}$ are the functions evaluated at the
parameter estimates.

Similar to calling the \code{systemfit} function, \code{nlsystemfit}
is called with a minimum of three arguments,

\code{
R> nlsystemfit( method, eqns, start )
}

where \code{method} is one of the following estimation methods: "OLS",
"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
to those described for \code{systemfit}, and \code{start} is a list of
starting values of the parameter estimates. The nlsystemfit function
relies on \code{nlm} to perform the minimization of the objective
functions and the \code{qr} set of functions. If the user does not
have a set of estimates for the initial parameters, it is suggested
using linearized forms in one of the linear methods, or simply using
\code{nls} to obtain estimates for the equations.

% The mandatory argument
% \code{eqns} is a list of the equations to estimate.  Each equation is
% a standard formula in \proglang{R}.  It starts with a dependent
% variable on the left hand side.  After a tilde ($~$) the regressors
% are listed  \footnote{For Details see the \proglang{R} help files to
%   \code{formula}}.

% The objective functions for the methods are:

% \begin{center}
% \begin{tabular}{|l|c|c|c|} \hline
%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
% \end{tabular}
% \end{center}

% where, $r$ is a column vector for the residuals for each equation,
% what is the difference to \epsilon?
% $X$ is matrix of the partial derivatives of the dependent variable 
% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
% is this correct?
% in the linear section X is used for the regressors, 
% thus we should use a different variable name here.
% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
% matrix of the instrument variables, and $I$ is an $n \times n $
% identity matrix and $S$ is the estimated variance-covariance matrix between the
% equations

% \begin{equation}
%   \label{eq:non-linear_varcov}
%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
% \end{equation}

% Do we actually have the variance-covariance function in place.

% The argument \code{method} is a string determining the estimation method.
% It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".

% \textbf{The residual variance-covariance matrix can be calculated in
%   different ways (section~\ref{sec:residcov}).  It should be
%   relatively easy to implement this also in nlsystemfit(),  e.g. the
%   function 'calcRCov' in systemfit() could be moved outside
%   systemfit() that it can be used also by nlsystemfit().}

% \textbf{You need to clear this up.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End:

### 2005-11-02 paper/trunk/systemfit_summary.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@574 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

What should we place in this section?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The \pkg{systemfit} package was originally developed to fit
simultaneous equations for forestry datasets. Forestry datasets
typically contain observations for many inexpensive observations
(diameter at breast height) and very few expensive observations such
as total tree height. While the package currently does not allow for
unbalanced datasets, data where the equations contain different
numbers of observations, future releases of the package will have that
feature implemented. We are still trying to figure out how develop the
code for instrumental variable methods (2SLS,3SLS) for the correct
estimation methods if possible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Additional Methods}\label{sec:summary_additional_methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Full Information Maximum Likelihood}

Describe, in a paragraph the process and give some reasons why you
might want this in the package.

\subsubsection{Generalized Method of Moments}

Describe, in a paragraph the process and give some reasons why you
might want this in the package.

>>>>

The \pkg{systemfit} package was originally developed to fit
simultaneous equations for forestry datasets. Forestry datasets
typically contain observations of many inexpensive observations (stem
diameter) and few expensive observations such as stem height and being
able the utilize as many observations as possible is advantageous. As
most software packages simply drop data points that do not contain
both observations, the \pkg{systemfit} package currently does not
allow for unbalanced datasets, but it is our intention to include this
capability for certain estimation methods in future releases.

In this article, we have described some of the basic features of the
\pkg{systemfit} package for estimation of linear systems of
equations. It has been tested on a variety of datasets and has
produced satisfactory for a few years. While more sophisticated tools
exist, the \pkg{systemfit} package will perform the basic fitting
methods. We wish to include more sophisticated estimation methods such
as full-information maximum likelihood and generalized methods of
moments.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End:
